### 2018-10-22 src/3_TimeSeries.tex
Adding all before the exam
correccion despues del examen

Figure \ref{fig:returnacf} shows the ACF of the returns of SPY stock where the
correlations are significant even for very long lags, this implies a long-memory
process.

 \begin{figure}[h]
 \centering
 \includegraphics[width=0.9\textwidth]{plots/spy_returns_acf.pdf}
 \caption{SPY returns ACF}
 \label{fig:returnacf}
\end{figure}
\item[Distribution] The distribution of returns is approximately symmetric and
has high kurtosis (i.e fat tails and a peaked centre compared with the normal
distribution). 
The returns distribution tails are larger than what is hypothesised by common
data generation process (generally normal distribution assumption). In the
markets, fat tails are an undesirable feature because of the additional risk
they imply.  However, distribution of returns whose were obtained from higher
frequencies looks more like a normal distribution.
In the figure \ref{fig:returndist} is shown the SPY returns distribution based
on daily dates from the period 1st July 1998 to 4th April 2013. The distribution
was compared against the normal distribution which clearly doesn't fit the data.  
 \begin{figure}[h]
 \centering
 \includegraphics[scale=0.5]{plots/spy_returns_dist.pdf}
 % spy_returns_dist.pdf: 504x504 pixel, 72dpi, 17.78x17.78 cm, bb=0 0 504 504
 \caption{SPY returns distribution}
 \label{fig:returndist}
\end{figure}
\item[Heterogeneity] despite the fact that financial returns are non-stationary,
stationary periods can be observed. Economist speaks in terms of a structural
break \cite{stock1994}, in machine learning this is known as drift which means
that the statistical properties of the target variable change over time
\cite{widmer1996}, \cite{tsymbal2004}.
\item[Non-Linearity] financial returns may be non-linear in mean and/or
non-linear in variance.
\item[Calendar effects] are also called seasonal effects and they are cyclical
anomalies in returns where the cycle is based on the calendar. Some of known
calendar effects are: intraday effect, weekend effect, Monday effect, intramonth
effect, the January effect and Holiday effect. The most important calendar
anomalies are the January effect and the weekend effect. 
\end{description}

@@

\begin{enumerate}[(a)]
\item \textbf{Deterministic trend:} Deterministic trend or trend stationary
processes have the following form:

\[
y_t = f(t) + \epsilon_{t} \quad ,
\]

\noindent where $t$ is the time trend and $\epsilon_{t}$ represents a stationary
error term (with mean 0 and variance $\sigma^2$) and $f(t)$ is a deterministic
function of time:
\begin{itemize}
\item If $f(t) = \alpha + \beta t$ we have a linear trend model which is widely used. 
\item If $f(t)= \alpha \exp^rt$ we have an exponential growth curve.
\item If $f(t) = c_1 + c_2t + c_3 t^2$ we have a quadratic trend model
\item If $f(t) = \frac{1}{k+\alpha \beta ^t }$ we have a logistic curve
\end{itemize}
\item \textbf{Stochastic trend:} Stochastic trend processes are also called unit
root or difference stationarity processes and have the following form:

\[
y_t = \mu + y_{t-1} + \epsilon_t
\]
\noindent where $\epsilon_t$ is a stationary process. When $\mu = 0 $ the
process is called pure random walk and when $\mu \neq 0$  the process is called
random walk with drift.

@@

A random walk with or without drift can be transformed to a stationary process
by differencing the time serie once. The disadvantage of differencing is that
the process loses one observation each time the time series is differentiated.

@@

In presence of spurious relationships the denominator of equation
\ref{eq:rsquared} becomes very large because a large weight is placed on
extreme observations on either side of the mean $\bar{y}$.

@@

Figure \ref{fig:spurious1} shows the time series $x_{1,t}$ and $x_{2,t}$ and
figure \ref{fig:spurious2} shows the regression between them and how they seem
to be related.
\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/spurious1}
  \caption[Random walks time series]{Two random walks time series $x_{1,t}$ and $x_{2,t}$}
  \label{fig:spurious1}
\end{figure}

@@

These two random walks, which are independent by construction, appear to be
related with a $R^2=0.883$. However, what it is happening is that they are
drifting in the same direction. Drift can be removed using returns or first
differences.  If after taking returns or first differences the regression fit
is still good we will say that the variables are cointegrated, this means that
both series may drift but their residuals will not. A more formal definition of
cointegration is given in the section \ref{sec:cointegration}.

@@

There are two statistics for cointegration under Johansen approach: the trace
(shown in equation \ref{eq:johansentrace}) and the maximum-eigenvalue statistic
(shown in equation \ref{eq:johansenmax}):

@@

\noindent where $r$ is the number of cointegration vectors and
$\hat]{\lambda}_i$ is the estimated value for the ith ordered eigenvalue.
$\lambda_{\text{trace}}$ null hypothesis is that the number of cointegration
vectors is less than or equal to $r$ against that there are more than $r$.
$\lambda_{\text{max}}$ has the null hypothesis that the number of cointegration
vectors is $r$ against an alternative of $r +1$.

@@

If we have two-dimensional process $\mathbf{y}_t$, $t=1,\dots,T$ by:

@@

Figure \ref{fig:cointex} shows the three time series examples.

@@

This example shows how cointegration vectors describes the stable relations
between the processes by linear relations that are more stationary than the
original process. Cointegration vectors can be obtained using the Johansen
method which gives with a certain probability the number of significant vectors.
Figures \ref{fig:cointvectors3}, \ref{fig:cointvectors4}  and
\ref{fig:cointvectors5}  shows the linear combinations obtained using
cointegration vectors given by Johansen method. The method confirm that only two
vectors are found and the third one given doesn't result in a I(0) process.
\vfill

@@

The $\text{ADF}_t$ statistic is based on the least squares estimates of equation \ref{eq:adftest}.

@@

\subsubsection{Realised volatility} is also known as historic volatility and
it is the actual variance in the price of a stock over time.
Realised volatility is measured in terms of the standard deviation
using the historical stock prices. It is commonly calculated based on
intraday price returns:
\begin{equation}
\label{eq:retintra}
r_{t,n}=100(\ln(p_{t,n}) - \ln(p_{t,n-1}))
\end{equation}
\noindent where $p_{t,n}$ is the price observed at day $t=1,\dots,T$ and
intraday sample $n=2,\dots,N$. Realised volatility is defined as:
\begin{equation}
\label{eq:rv}
    \hat{\sigma}(t) = \sum_{n=1}^N r_{t,n}^2 \, , 
\end{equation}
\noindent where $N$ is the number of intraday samples and $T$ is the
number of days. 
In order to include overnight returns, Hansen and
Lunde~\cite{hansen+lunde2005} introduced a scaling version of
realised volatility using the following definitions:
\begin{eqnarray}
r_{t}&=&100(\ln(p_{t,N}) - \ln(p_{t-1,N})) \label{eq: ret 1} \\
\bar{\rho}(t) &=& \sum_{t=1}^T r_{t}^2  \label{eq: ret 2} \, .
\end{eqnarray}
\noindent where equation~(\ref{eq: ret 1}) represents overnight 
returns and the volatility as 
equation~(\ref{eq: ret 2}), where $p_{t,N}$ is the last intraday 
sample at day $t$. The scaled realised volatility $\rho(t)$ is 
defined as:
\begin{eqnarray}
\label{eq:srv}
\rho(t) = \gamma \hat{\rho}(t) \, , \qquad & \qquad \gamma = \displaystyle \frac{\bar{\rho}(t)}{\displaystyle\sum_{t=1}^T \hat{\rho}(t)}
\end{eqnarray}
Realised volatility has also been defined as the absolute value return or as
the mean of the sum of intraday squared returns at short intervals of time. The
majority of research carried out in the literature obtain the daily volatility
as the daily squared returns as is shown in equation~(\ref{eq: ret 2}).
However, it has been proven that this measurement noise is too high for
observing the true volatility process~\cite{andersen+bollerslev1998}. Hansen
and Lunde~\cite{hansen+lunde2006} stated that the use of a noisy proxy could
result in an inferior model being chosen as the best one. The realised
volatility, as calculated by the cumulative sum of squared intraday returns and
shown in equation (\ref{eq:rv}), is less noisy and doesn't lead to choosing an
inferior model.   
\subsubsection{Implied volatility} not only can be extracted from returns
but it can also be derived from option or future pricing models.  The
volatility obtained corresponds to the market's prediction of future
volatility. In finance, an option is a derivative, that is, a contract which
gives the owner the right, but not the obligation to buy or sell an underlying
asset at a given price called strike price. An option can be executed at any
time before an expiration date previously defined no matter what price the
underlying asset has. For example, the Black-Scholes model~\cite{black1973}
determines the fair option value based on stock price, strike price, time to
option expiration, the interest rate and volatility. These are known or can be
easily obtained from the market, excepting by volatility which must be
estimated. However, rather than assuming a volatility a priori and computing
option prices from it, the model can be used to estimate volatility at given
prices, time to expiration and strike price. This obtained volatility is called
the implied volatility of an option. Additionally, some models obtain implied
volatility from futures (other derivative from prices). For instance, the
Barone-Adesi and Whaley futures option model~\cite{baroneetal1987} is also used
to determine future volatilities~\cite{hamidetal2004}. Higher implied
volatility is indicative of greater price fluctuation in either direction.
Implied volatility is found by determining the value which makes theoretical
prices equal to market prices. In this way volatility is ``implied'' by the
current market price of the stock.

@@

\noindent Therefore, the integral form of the stock price model shown in
equation (\ref{eq:stockprice}) is:

@@

The random walk model, despite its simplicity, is still difficult to outperform
for standard econometric forecasting models \cite{lo2011}. The random walk model
is defined as:

\begin{equation}
\mathbf{y}_t = \mathbf{y}_{t-1} + \epsilon_{t}
\label{rwmodel}
\end{equation}

@@

The vector autoregressive VAR($p$) model \cite{sims1980} is one of the most easy
to use, successful and flexible models for the analysis of multivariate time
series. It is a natural extension of the univariate autoregressive (AR) model.
The VAR model has proven to be useful for describing the dynamic behaviour of
economic and financial time series.  VAR is a general framework describing the
behaviour of a set of $l$ endogenous variables as a linear combination of their
last $p$ values, where $l,p\in\mathbb{N}$.  In our case, each one of these $l$
variables is a scalar time series $y_{i,t}$, $i=1,\dots,l$, and we represent
them all together at time $t$ by the the vector time series:
\begin{equation*}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} & y_{2,t} & \dots & y_{l,t} \end{bmatrix}^\top.
\end{equation*}
\noindent
Notice that the vector $\mathbf{y}_t$ is assumed to be $l$-dimensional.

@@

Notice that, regarding our notation of the cointegration section
(\ref{sec:cointegration}), we have here 
$\mathbf{y}_t^0 = \mathbf{y}_t$,
$\mathbf{y}_t^\nu = \mathbf{y}_{t-\nu}$ and
the $i$-th component of the vector time series $\mathbf{y}_t^\nu$
is the scalar time series $y_{i,t}^\nu$, where $\nu=1,\dots,p$ and
$i=1,\dots,l$.
Transposing each equation of the system (\ref{eq:var}) we can write
the VAR($p$) model in block-matrix form as:
\begin{equation}\label{eq:simplevareq}
\mathbf{Y} = \mathbf{A} \mathbf{X} + \mathbf{E} \, , 
\end{equation}
%%
\noindent where:
%%
\begin{alignat*}{2}
\mathbf{Y}
&= \begin{bmatrix}
   \mathbf{y}_{p+1}^\top \\
   \mathbf{y}_{p+2}^\top \\
   \vdots \\
   \mathbf{y}_N^\top
   \end{bmatrix}_{(N-p)\times l}
&\quad
\mathbf{A}
&= \begin{pmat}[{...|}]
   \mathbf{y}_p^\top & \mathbf{y}_{p-1}^\top & \dots 
                    & \mathbf{y}_1^\top & 1 \cr
   \mathbf{y}_{p+1}^\top & \mathbf{y}_p^\top & \dots
                       & \mathbf{y}_2^\top & 1 \cr
   \vdots & \vdots & \ddots & \vdots & \vdots \cr
   \mathbf{y}_{N-1}^\top & \mathbf{y}_{N-2}^\top & \dots 
                       & \mathbf{y}_{N-p}^\top & 1 \cr
   \end{pmat}_{(N-p)\times(pl+1)} \\
\mathbf{X}
&= \begin{bmatrix}
   \boldsymbol{\Phi}_1^\top \\
   \boldsymbol{\Phi}_2^\top \\
   \vdots \\
   \boldsymbol{\Phi}_p^\top \\
   \mathbf{c}^\top
   \end{bmatrix}_{(pl+1)\times l}
&\quad
\mathbf{E}
&= \begin{bmatrix}
   \boldsymbol{\epsilon}_{p+1}^\top \\
   \boldsymbol{\epsilon}_{p+2}^\top \\
   \vdots \\
   \boldsymbol{\epsilon}_N^\top \\
   \end{bmatrix}_{(N-p)\times l}
\end{alignat*}
Taking into account the error term $\mathbf{E}$, equation \ref{eq:simplevareq}
can be solved with respect to $\mathbf{X}$ using the ordinary least
squares estimation (see section \ref{sec:OLS}).

@@

These difficulties were illustrated by Granger and Newbold in 1974
\cite{granger1974} when they introduced the concept of spurious regressions (see
section \ref{sec:spurious}). In 1982, Nelson and Plosser \cite{nelson1982}
showed that unit roots might be present in a wide variety of macroeconomic
series in levels or logarithms. This finding made the unit root testing very
popular. Aditionally, the implication that variables should be rendered
stationary by differencing before they are included in an econometric model was
generalised.  Further theoretical developments by Granger and Engle in 1987
\cite{engle87} raised the possibility that two or more integrated, non
stationary time series might be cointegrated, so that some linear combination of
these series could be stationary even though each series were not.

@@

VECM  is obtained re-writing equation \ref{eq:var} in terms of the new
variable $\Delta\mathbf{y}_t=\mathbf{y}_t-\mathbf{y}_{t-1}$.
The VECM model, expressed in terms those differences, takes the form:
\begin{equation}\label{eq:vec}
\Delta \mathbf{y}_t 
= \boldsymbol{\Omega}\,\mathbf{y}_{t-1}
  + \sum_{i=1}^{p-1} \boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
  + \mathbf{c} + \boldsymbol{\epsilon}_t\,,
\end{equation}
\noindent
where the coefficients matrices $\boldsymbol{\Phi}_i^*$ and 
$\boldsymbol{\Omega}$, expressed in terms of the matrices
$\boldsymbol{\Phi}_i$ of the model VAR shown in equation \ref{eq:var}, are:
\begin{align*}
\boldsymbol{\Phi}_i^* 
&:= -\sum_{j=i+1}^{p}\boldsymbol{\Phi}_j\,, \\
\boldsymbol{\Omega}
&:= -\left( \mathbb{I} - \boldsymbol{\Phi}_1 - \dots 
    - \boldsymbol{\phi}_p \right)\,. 
\end{align*}
The following well known properties of the matrix $\boldsymbol{\Omega}$
\cite{johansen1995} will be useful in the sequel:
\begin{itemize}
\item
If $\boldsymbol{\Omega} = \mathbf{0}$, there is no cointegration.
\item 
If $rank(\boldsymbol{\Omega})=l$, i.e., if $\boldsymbol{\Omega}$ has
full rank, then the time series are not I(1) but stationary.
\item
If $rank(\boldsymbol{\Omega})=r$, $0<r<l$, then there is cointegration
and the matrix $\boldsymbol{\Omega}$ can be expressed as
$\boldsymbol{\Omega}=\boldsymbol{\alpha\beta}^\top$, where $\boldsymbol{\alpha}$
and $\boldsymbol{\beta}$ are
$l\times r$ matrices and
$\text{rank}(\boldsymbol{\alpha})=\text{rank}(\boldsymbol{\beta})=r$.
\item
The columns of $\boldsymbol{\beta}$ contains the cointegration vectors and the rows of
$\boldsymbol{\alpha}$ correspond with the adjusted vectors. 
$\boldsymbol{\beta}$ is obtained by Johansen procedure~\cite{johansen1988},
whereas $\boldsymbol{\alpha}$ has to be determined as a variable in the VECM.
\end{itemize}
It is worth noticing that the factorisation of the matrix
$\boldsymbol\Omega$ is not unique, since for any $r \times r$
nonsingular matrix $\mathbf{H}$, $\boldsymbol{\alpha}^*:=\boldsymbol{\alpha}\mathbf{H}$,
and $\boldsymbol{\beta}^*=\boldsymbol{\beta}(\mathbf{H}^{-1})^\top$ we have
$\boldsymbol{\alpha\beta}^\top=\boldsymbol{\alpha}^*(\boldsymbol{\beta}^*)^\top$.
If cointegration exists, then equation (\ref{eq:vec}) can be written
as follows:
\begin{equation}\label{eq:vecfull}
\Delta\mathbf{y}_t 
= \boldsymbol{\alpha\beta}^\top\mathbf{y}_{t-1} 
  + \sum_{i=1}^{p-1}\boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
  + \mathbf{c} + \boldsymbol{\epsilon}_t\,,
\end{equation}
\noindent
which is a VAR model but for time series differences.

Transposing each equation of the system (\ref{eq:vecfull}) we can write
the VECM($p$) model in block-matrix form as:
\begin{equation}\label{eq:simpleveceq}
\mathbf{B} = 
\mathbf{A} \mathbf{X} + 
\mathbf{E} \, , 
\end{equation}

\noindent where $\mathbf{Y}$ dimension is $((N-p)\times l)$, $\mathbf{A}$
dimension is $((N-p)\times(r+(p-1)l +1))$, $\mathbf{X}$ dimension is $((r+(p-1)l
+1)\times l)$ and $\mathbf{E}$ dimension is $((N-p)\times l)$:

@@

More details can be found in section \ref{app:pscsvd}.

>>>>

Figure~\ref{fig:returnacf} shows the ACF of the returns of SPY stock where the
correlations are significant even for very long lags, this implies a long-memory
process.

 \begin{figure}[h]
 \centering
 \includegraphics[width=0.9\textwidth]{plots/spy_returns_acf.pdf}
 \caption{SPY returns ACF}
 \label{fig:returnacf}
\end{figure}
\item[Distribution] The distribution of returns is approximately symmetric and
has high kurtosis (i.e fat tails and a peaked centre compared with the normal
distribution). 
The returns distribution tails are larger than what is hypothesised by common
data generation process (generally normal distribution assumption). In the
markets, fat tails are an undesirable feature because of the additional risk
they imply.  However, distribution of returns whose were obtained from higher
frequencies looks more like a normal distribution.
In figure~\ref{fig:returndist} is shown the SPY returns distribution based
on daily dates from the period 1st July 1998 to 4th April 2013. The distribution
was compared against the normal distribution which clearly doesn't fit the data.  
 \begin{figure}[h]
 \centering
 \includegraphics[scale=0.5]{plots/spy_returns_dist.pdf}
 % spy_returns_dist.pdf: 504x504 pixel, 72dpi, 17.78x17.78 cm, bb=0 0 504 504
 \caption{SPY returns distribution}
 \label{fig:returndist}
\end{figure}
\item[Heterogeneity] despite the fact that financial returns are non-stationary,
stationary periods can be observed. Economist speaks in terms of a structural
break \cite{stock1994}, in machine learning this is known as drift which means
that the statistical properties of the target variable change over time
\cite{widmer1996}, \cite{tsymbal2004}.
\item[Non-Linearity] financial returns may be non-linear in mean and/or
non-linear in variance.
\item[Calendar effects] are also called seasonal effects and they are cyclical
anomalies in returns where the cycle is based on the calendar. Some of known
calendar effects are: intraday effect, weekend effect, Monday effect, intramonth
effect, the January effect and Holiday effect. The most important calendar
anomalies are the January effect and the weekend effect. 
\end{description}

@@

\begin{enumerate}[(a)]
\item \textbf{Deterministic trend:} Deterministic trend or trend stationary
processes have the following form:
%
\[
y_t = f(t) + \epsilon_{t} \quad ,
\]
%
\noindent where $t$ is the time trend and $\epsilon_{t}$ represents a stationary
error term (with mean 0 and variance $\sigma^2$) and $f(t)$ is a deterministic
function of time:
\begin{itemize}
\item If $f(t) = \alpha + \beta t$ we have a linear trend model which is widely used. 
\item If $f(t)= \alpha \exp^{rt}$ we have an exponential growth curve.
\item If $f(t) = c_1 + c_2t + c_3 t^2$ we have a quadratic trend model
\item If $f(t) = \displaystyle\frac{1}{k+\alpha \beta ^t }$ we have a logistic curve
\end{itemize}
\item \textbf{Stochastic trend:} Stochastic trend processes are also called unit
root or difference stationarity processes and have the following form:
\[
y_t = \mu + y_{t-1} + \epsilon_t
\]
\noindent where $\epsilon_t$ is a stationary process. When $\mu = 0 $ the
process is called pure random walk and when $\mu \neq 0$  the process is called
random walk with drift.

@@

A random walk with or without drift can be transformed to a stationary process
by differencing the time series once. The disadvantage of differencing is that
the process loses one observation each time the time series is differentiated.

@@

In presence of spurious relationships the denominator of equation
\eqref{eq:rsquared} becomes very large because a large weight is placed on
extreme observations on either side of the mean $\bar{y}$.

@@

Figure~\ref{fig:spurious1} shows the time series $x_{1,t}$ and $x_{2,t}$ and
figure~\ref{fig:spurious2} shows the regression between them and how they seem
to be related.
\begin{figure}[!h]
  \centering
  \includegraphics[width=0.8\textwidth]{img/spurious1}
  \caption[Random walks time series]{Two random walks time series $x_{1,t}$ and $x_{2,t}$}
  \label{fig:spurious1}
\end{figure}

@@

These two random walks, which are independent by construction, appear to be
related with a $R^2=0.883$. However, what it is happening is that they are
drifting in the same direction. Drift can be removed using returns or first
differences.  If after taking returns or first differences the regression fit
is still good we will say that the variables are cointegrated, this means that
both series may drift but their residuals will not. A more formal definition of
cointegration is given in the section~\ref{sec:cointegration}.

@@

There are two statistics for cointegration under Johansen approach: the trace
(shown in equation~\eqref{eq:johansentrace}) and the maximum-eigenvalue statistic
(shown in equation~\eqref{eq:johansenmax}):

@@

\noindent where $r$ is the number of cointegration vectors, $T$ is the number of observations and
$\hat{\lambda}_i$ is the estimated value for the ith ordered eigenvalue.
$\lambda_{\text{trace}}$ null hypothesis is that the number of cointegration
vectors is less than or equal to $r$ against that there are more than $r$.
$\lambda_{\text{max}}$ has the null hypothesis that the number of cointegration
vectors is $r$ against an alternative of $r +1$.

@@

If we have a two-dimensional process $\mathbf{y}_t$, $t=1,\dots,T$ by:

@@

Figure~\ref{fig:cointex} shows the three time series examples.

@@

This example shows how cointegration vectors describes the stable relations
between the processes by linear relations that are more stationary than the
original process. Cointegration vectors can be obtained using the Johansen
method which gives with a certain probability the number of significant vectors.
Figures~\ref{fig:cointvectors3},~\ref{fig:cointvectors4}  and
\eqref{fig:cointvectors5}  shows the linear combinations obtained using
cointegration vectors given by Johansen method. The method confirm that only two
vectors are found and the third one given doesn't result in a I(0) process.
\vfill

@@

The $\text{ADF}_t$ statistic is based on the least squares estimates of equation~\eqref{eq:adftest}.

@@

\subsubsection{Realised volatility} is also known as historic volatility and
it is the actual variance in the price of a stock over time.
Realised volatility is measured in terms of the standard deviation
using the historical stock prices. It is commonly calculated based on
intraday price returns:
\begin{equation}
\label{eq:retintra}
r_{t,n}=100(\ln(p_{t,n}) - \ln(p_{t,n-1}))
\end{equation}
\noindent where $p_{t,n}$ is the price observed at day $t=1,\dots,T$ and
intraday sample $n=2,\dots,N$. Realised volatility is defined as:
\begin{equation}
\label{eq:rv}
    \hat{\sigma}(t) = \sum_{n=1}^N r_{t,n}^2 \, , 
\end{equation}
\noindent where $N$ is the number of intraday samples and $T$ is the
number of days. 
In order to include overnight returns, Hansen and
Lunde~\cite{hansen+lunde2005} introduced a scaling version of
realised volatility using the following definitions:
\begin{eqnarray}
r_{t}&=&100(\ln(p_{t,N}) - \ln(p_{t-1,N})) \label{eq:ret1} \\
\bar{\rho}(t) &=& \sum_{t=1}^T r_{t}^2  \label{eq:ret2} \, .
\end{eqnarray}
\noindent where equation~\eqref{eq:ret1} represents overnight 
returns and the volatility as 
equation~\eqref{eq:ret2}, where $p_{t,N}$ is the last intraday 
sample at day $t$. The scaled realised volatility $\rho(t)$ is 
defined as:
\begin{eqnarray}
\label{eq:srv}
\rho(t) = \gamma \hat{\rho}(t) \, , \qquad & \qquad \gamma = \displaystyle \frac{\bar{\rho}(t)}{\displaystyle\sum_{t=1}^T \hat{\rho}(t)}
\end{eqnarray}
Realised volatility has also been defined as the absolute value return or as
the mean of the sum of intraday squared returns at short intervals of time. The
majority of research carried out in the literature obtain the daily volatility
as the daily squared returns as is shown in equation~\eqref{eq:ret2}.
However, it has been proven that this measurement noise is too high for
observing the true volatility process~\cite{andersen+bollerslev1998}. Hansen
and Lunde~\cite{hansen+lunde2006} stated that the use of a noisy proxy could
result in an inferior model being chosen as the best one. The realised
volatility, as calculated by the cumulative sum of squared intraday returns and
shown in equation~\eqref{eq:rv}, is less noisy and doesn't lead to choosing an
inferior model.   
\subsubsection{Implied volatility} not only can be extracted from returns
but it can also be derived from option or future pricing models.  The
volatility obtained corresponds to the market's prediction of future
volatility. In finance, an option is a derivative, that is, a contract which
gives the owner the right, but not the obligation to buy or sell an underlying
asset at a given price called strike price. An option can be executed at any
time before an expiration date previously defined no matter what price the
underlying asset has. For example, the Black-Scholes model~\cite{black1973}
determines the fair option value based on stock price, strike price, time to
option expiration, the interest rate and volatility. These are known or can be
easily obtained from the market, excepting by volatility which must be
estimated. However, rather than assuming a volatility a priori and computing
option prices from it, the model can be used to estimate volatility at given
prices, time to expiration and strike price. This obtained volatility is called
the implied volatility of an option. Additionally, some models obtain implied
volatility from futures (other derivative from prices). For instance, the
Barone-Adesi and Whaley futures option model~\cite{baroneetal1987} is also used
to determine future volatilities~\cite{hamidetal2004}. Higher implied
volatility is indicative of greater price fluctuation in either direction.
Implied volatility is found by determining the value which makes theoretical
prices equal to market prices. In this way volatility is ``implied'' by the
current market price of the stock.

@@

\noindent Therefore, the integral form of the stock price model shown in
equation~\eqref{eq:stockprice} is:

@@

The random walk model, despite its simplicity, is still difficult to outperform
for standard econometric forecasting models \cite{lo2011}. The random walk model
is defined as:
%
\begin{equation}
\mathbf{y}_t = \mathbf{y}_{t-1} + \epsilon_{t}
\label{rwmodel}
\end{equation}

@@

There are several multivariate models widely used such as the multivariate ARCH model \cite{bollerslev1990modelling} and variations like the DCC-GARCH \cite{engle2002dynamic}, the BEKK model \cite{10.2307/3532933} and the VAR model  \cite{sims1980}. 
The vector autoregressive VAR($p$) model derives the vector correction model VECM, which studies the cointegration between time series and also is one of the most easy
to use and flexible models for the analysis of multivariate time
series.

The VAR model is a natural extension of the univariate autoregressive (AR) model.
The VAR model has proven to be useful for describing the dynamic behaviour of
economic and financial time series. VAR is a general framework describing the
behaviour of a set of $l$ endogenous variables as a linear combination of their
last $p$ values, where $l,p\in\mathbb{N}$.  In our case, each one of these $l$
variables is a scalar time series $y_{i,t}$, $i=1,\dots,l$, and we represent
them all together at time $t$ by the the vector time series:
\begin{equation*}
\label{eq:variables}
\mathbf{y}_t = 
\begin{bmatrix} y_{1,t} & y_{2,t} & \dots & y_{l,t} \end{bmatrix}^\top.
\end{equation*}
\noindent
Notice that the vector $\mathbf{y}_t$ is assumed to be $l$-dimensional.

@@

Notice that, regarding our notation of the cointegration section
\eqref{sec:cointegration}, we have here 
$\mathbf{y}_t^0 = \mathbf{y}_t$,
$\mathbf{y}_t^\nu = \mathbf{y}_{t-\nu}$ and
the $i$-th component of the vector time series $\mathbf{y}_t^\nu$
is the scalar time series $y_{i,t}^\nu$, where $\nu=1,\dots,p$ and
$i=1,\dots,l$.
Transposing each equation of the system~\eqref{eq:var} we can write
the VAR($p$) model in block-matrix form as:
\begin{equation}\label{eq:simplevareq}
\mathbf{Y} = \mathbf{A} \mathbf{X} + \mathbf{E} \, , 
\end{equation}
%%
\noindent where:
%%
\begin{alignat*}{2}
\mathbf{Y}
&= \begin{bmatrix}
   \mathbf{y}_{p+1}^\top \\
   \mathbf{y}_{p+2}^\top \\
   \vdots \\
   \mathbf{y}_N^\top
   \end{bmatrix}_{(N-p)\times l}
&\quad
\mathbf{A}
&= \begin{pmat}[{...|}]
   \mathbf{y}_p^\top & \mathbf{y}_{p-1}^\top & \dots 
                    & \mathbf{y}_1^\top & 1 \cr
   \mathbf{y}_{p+1}^\top & \mathbf{y}_p^\top & \dots
                       & \mathbf{y}_2^\top & 1 \cr
   \vdots & \vdots & \ddots & \vdots & \vdots \cr
   \mathbf{y}_{N-1}^\top & \mathbf{y}_{N-2}^\top & \dots 
                       & \mathbf{y}_{N-p}^\top & 1 \cr
   \end{pmat}_{(N-p)\times(pl+1)} \\
\mathbf{X}
&= \begin{bmatrix}
   \boldsymbol{\Phi}_1^\top \\
   \boldsymbol{\Phi}_2^\top \\
   \vdots \\
   \boldsymbol{\Phi}_p^\top \\
   \mathbf{c}^\top
   \end{bmatrix}_{(pl+1)\times l}
&\quad
\mathbf{E}
&= \begin{bmatrix}
   \boldsymbol{\epsilon}_{p+1}^\top \\
   \boldsymbol{\epsilon}_{p+2}^\top \\
   \vdots \\
   \boldsymbol{\epsilon}_N^\top \\
   \end{bmatrix}_{(N-p)\times l}
\end{alignat*}
Taking into account the error term $\mathbf{E}$, equation~\eqref{eq:simplevareq}
can be solved with respect to $\mathbf{X}$ using the ordinary least
squares estimation (see section~\ref{sec:OLS}).

@@

These difficulties were illustrated by Granger and Newbold in 1974
\cite{granger1974} when they introduced the concept of spurious regressions (see
section~\ref{sec:spurious}). In 1982, Nelson and Plosser \cite{nelson1982}
showed that unit roots might be present in a wide variety of macroeconomic
series in levels or logarithms. This finding made the unit root testing very
popular. Aditionally, the implication that variables should be rendered
stationary by differencing before they are included in an econometric model was
generalised.  Further theoretical developments by Granger and Engle in 1987
\cite{engle87} raised the possibility that two or more integrated, non
stationary time series might be cointegrated, so that some linear combination of
these series could be stationary even though each series were not.

@@

VECM  is obtained re-writing equation~\eqref{eq:var} in terms of the new
variable $\Delta\mathbf{y}_t=\mathbf{y}_t-\mathbf{y}_{t-1}$.
The VECM model, expressed in terms those differences, takes the form:
\begin{equation}\label{eq:vec}
\Delta \mathbf{y}_t 
= \boldsymbol{\Omega}\,\mathbf{y}_{t-1}
  + \sum_{i=1}^{p-1} \boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
  + \mathbf{c} + \boldsymbol{\epsilon}_t\,,
\end{equation}
\noindent
where the coefficients matrices $\boldsymbol{\Phi}_i^*$ and 
$\boldsymbol{\Omega}$, expressed in terms of the matrices
$\boldsymbol{\Phi}_i$ of the model VAR shown in equation~\eqref{eq:var}, are:
\begin{align*}
\boldsymbol{\Phi}_i^* 
&:= -\sum_{j=i+1}^{p}\boldsymbol{\Phi}_j\,, \\
\boldsymbol{\Omega}
&:= -\left( \mathbb{I} - \boldsymbol{\Phi}_1 - \dots 
    - \boldsymbol{\Phi}_p \right)\,. 
\end{align*}
The following well known properties of the matrix $\boldsymbol{\Omega}$
\cite{johansen1995} will be useful in the sequel:
\begin{itemize}
\item
If $\boldsymbol{\Omega} = \mathbf{0}$, there is no cointegration.
\item 
If $rank(\boldsymbol{\Omega})=l$, i.e., if $\boldsymbol{\Omega}$ has
full rank, then the time series are not I(1) but stationary.
\item
If $rank(\boldsymbol{\Omega})=r$, $0<r<l$, then there is cointegration
and the matrix $\boldsymbol{\Omega}$ can be expressed as
$\boldsymbol{\Omega}=\boldsymbol{\alpha\beta}^\top$, where $\boldsymbol{\alpha}$
and $\boldsymbol{\beta}$ are
$l\times r$ matrices and
$\text{rank}(\boldsymbol{\alpha})=\text{rank}(\boldsymbol{\beta})=r$.
\item
The columns of $\boldsymbol{\beta}$ contains the cointegration vectors and the rows of
$\boldsymbol{\alpha}$ correspond with the adjusted vectors. 
$\boldsymbol{\beta}$ is obtained by Johansen procedure~\cite{johansen1988},
whereas $\boldsymbol{\alpha}$ has to be determined as a variable in the VECM.
\end{itemize}
It is worth noticing that the factorisation of the matrix
$\boldsymbol\Omega$ is not unique, since for any $r \times r$
nonsingular matrix $\mathbf{H}$, $\boldsymbol{\alpha}^*:=\boldsymbol{\alpha}\mathbf{H}$,
and $\boldsymbol{\beta}^*=\boldsymbol{\beta}(\mathbf{H}^{-1})^\top$ we have
$\boldsymbol{\alpha\beta}^\top=\boldsymbol{\alpha}^*(\boldsymbol{\beta}^*)^\top$.
If cointegration exists, then equation~\eqref{eq:vec} can be written
as follows:
\begin{equation}\label{eq:vecfull}
\Delta\mathbf{y}_t 
= \boldsymbol{\alpha\beta}^\top\mathbf{y}_{t-1} 
  + \sum_{i=1}^{p-1}\boldsymbol{\Phi}_i^*\,\Delta\mathbf{y}_{t-i}
  + \mathbf{c} + \boldsymbol{\epsilon}_t\,,
\end{equation}
\noindent
which is a VAR model but for time series differences.

Transposing each equation of the system (\eqref{eq:vecfull}) we can write
the VECM($p$) model in block-matrix form as:
\begin{equation}\label{eq:simpleveceq}
\mathbf{B} = 
\mathbf{A} \mathbf{X} + 
\mathbf{E} \, , 
\end{equation}

\noindent where $\mathbf{B}$ dimension is $((N-p)\times l)$, $\mathbf{A}$
dimension is $((N-p)\times(r+(p-1)l +1))$, $\mathbf{X}$ dimension is $((r+(p-1)l
+1)\times l)$ and $\mathbf{E}$ dimension is $((N-p)\times l)$:

@@

More details can be found in section \eqref{app:pscsvd}.
